---
title: "P8105 Homework 3"
author: "Junxian Chen (jc5314)"
date: "10/6/2019"
output: github_document
---

# Problem 1

```{r, message = FALSE}
library(tidyverse)
library(p8105.datasets)
data("instacart")
```

**"write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations."**

The dataset contains `r nrow(instacart)` observations and `r ncol(instacart)` variables.

* There are `r length(unique(pull(instacart, aisle)))` aisles in total here, and the aisles that the most items ordered from is '`r names(sort(table(pull(instacart, aisle)),decreasing=TRUE))[1]`'.

* A plot that shows the number of items ordered in each aisle, limiting to aisles with more than 10000 items ordered：

```{r, fig.width = 8}
aisle_plot = 
  instacart %>% 
  count(aisle, name = 'number_of_items_ordered') %>%
  filter(number_of_items_ordered > 10000) %>%
  arrange(number_of_items_ordered) %>%
  ggplot(aes(x = number_of_items_ordered, y = aisle)) + 
    geom_point() +
    xlab('Number of items ordered') +
    ylab('Aisle name') +
    labs(title = 'Number of items ordered in each aisle with more than 10000 items ordered')
  
# show the plot

aisle_plot
```

* A table showing the three most popular items in each of the aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits` (numbers of times each item is ordered are also included):

```{r}
pop_item = 
  instacart %>% 
  filter(
    aisle %in% c('baking ingredients', 'dog food care', 'packaged vegetables fruits')
  ) %>% 
  group_by(aisle) %>% 
  count(product_name, name = 'number_of_ordered_time') %>% 
  top_n(n = 5, wt = number_of_ordered_time) %>% 
  knitr::kable(col.names = c('Aisle name', 'Top 5 popular items on this aisle', 'Number of ordered time'),
               caption = 'Table 1: The 3 most popular items in the aisles and their numbers of time being ordered.')

# show the table

pop_item
```

* A table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week:

```{r}
table_p1 = 
  instacart %>% 
    filter(
    product_name %in% c('Pink Lady Apples', 'Coffee Ice Cream')
  ) %>% 
  group_by(order_dow) %>% 
  mutate(mean_hour = mean(order_hour_of_day)) %>% 
  select(product_name, order_dow, mean_hour) %>% 
  arrange(order_dow) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour,
    values_fn = list(mean_hour = mean)
  ) %>% 
  knitr::kable(col.names = c('Product Name', 'Sunday', 'Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday'),
               caption = 'Table 2: Mean hour of the day at which the product was ordered on each day of the week.') 

# show the table

table_p1
```

# Problem 2

